\protect A Web of Human Knowledge: Unravelling Wikipedia's First Link Network

### 2015-11-27 writeup/2015-11-wikipedia-network.tex
fixed ndegree

Our goal is to gain insight into how the information on Wikipedia is organized and related
by studying the structure of the FLN.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of first links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, ndegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

%\green{TODO:
%rewrite \\
%introduce metric\\
%find two or three salient points to discuss in each section \\
%tie into bigger picture
%\begin{itemize}
%    \item reorder so paths and flow are described first then funnels and visits
%    \item Basins expand explanation and break up into 2 paragraphs
%    \item add more captions to figures
%\end{itemize}
%}
We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.

@@

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by ndegree}
  }
  We rank each article by the number of direct first links to the article (ndegree). 
  \label{fig:ndegree list}

\end{figure*}
The highest ranking article by ndegree is 
"United States" with 80,249 direct first links. Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands."

"Philosophy" and other philosophical concepts with many traversal visits
are not high-ranking by ndegree.
"Philosophy" has an ndegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy," the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argued in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}

\end{figure*}
As a distribution, we examine the degree of the articles in the FLN with at least
one direct first link. The average ndegree 
is 3.6 direct links with a standard deviation of 89.5 links.
Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
9. Overall the ndegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.

@@

Measuring only traversal visits and ndegree is limiting as it does not distinguish whether a particular article in a cycle 
is a funnel, directing many more paths inside the cycle than others. 
To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
so that accumulation does not flow to all articles in a cycle.
The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
from an article directly funneling many paths.

>>>>

Our goal is to gain insight into how the information on Wikipedia is organized and related
by studying the structure of the FLN.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of first links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, indegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

\green{TODO:
rewrite \\
introduce metric\\
find two or three salient points to discuss in each section \\
tie into bigger picture
\begin{itemize}
    \item reorder so paths and flow are described first then funnels and visits
    \item Basins expand explanation and break up into 2 paragraphs
    \item add more captions to figures
\end{itemize}
}

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by indegree}
  }
  We rank each article by the number of direct first links to the article (indegree). 
  \label{fig:ndegree list}

\end{figure*}
The highest ranking article by indegree is 
"United States" with 80,249 direct first links. Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands."

"Philosophy" and other philosophical concepts with many traversal visits
are not high-ranking by indegree.
"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}
\end{figure*}

As a distribution, we examine the degree of the articles in the FLN with at least
one direct first link. The average indegree 
is 3.6 direct links with a standard deviation of 89.5 links.
Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
9. Overall the indegree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.

@@

We followed every possible path through the network, taking 232 million steps along the way to measure the accumulated number of traversal visits.

@@

@@

Measuring only traversal visits and indegree is limiting as it does not distinguish whether a particular article in a cycle 
is a funnel, directing many more paths inside the cycle than others. 
To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
so that accumulation does not flow to all articles in a cycle.
The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
from an article directly funneling many paths.
